Supplementary methods

### 2021-06-30 text/supplemental_methods.Rmd

$$A = A_{0} A_{cl} + A_r (A'-20)$$ Where $A_0$ is the tissue specific effective age
parameter (expressed in years), $A_{cl}$ is a dimensionless multiplier for the
given class, and $A_r$ is the dimensionless age rate which determines how actual
age $A'$ affects effective age A using a linear perturbation.

@@

The product $A_0 A_{cl}$ is obviously symmetric between the two factors. To
disambiguate the meaning of the two, the prior for $A_0$ ranges from about 40 to
100 years, whereas the prior distribution of $A_{cl}$ has peak
probability at 1 and a relatively narrow width, due to the fact that it is a
tissue specific multiplicative modifying factor. To understand the roles of
these priors refer to (**SM3 Fig**).

Using this model, we split the data by Male/Female sex, and by White _vs._
Black race as these were the main well-identified categories available to us in
the dataset. Sex is known to be a risk factor for death, with males dying at
slightly higher rates for all ages of interest in our dataset. Black _vs._ White
race is an important category due to differences in access to care and other
socioeconomic and education related factors. However, due to the low number of
Black patient tumors, we use ethnicity only to identify the baseline risk
function, while the tissue- and cluster-specific multipliers are not ethnicity
specific. There were too few tumors from other races to obtain good inference so
we excluded them from our analysis.

>>>>

$$A = A_{0} A_{cl} + A_r (A'-20)$$ Where $A_0$ is the tissue specific effective
age parameter (expressed in years), $A_{cl}$ is a dimensionless multiplier for
the given class, and $A_r$ is the dimensionless age rate which determines how
actual age $A'$ affects effective age A using a linear perturbation. For a
non-cancer patient, $A_0=20$, $A_{cl}=1$, and $A_r=1$ are the relevant
benchmarks.

@@

The product $A_0 A_{cl}$ is obviously symmetric between the two factors. To
disambiguate the meaning of the two, the prior for $A_0$ has high probability
range from about 40 to 100 years, whereas the prior distribution of $A_{cl}$ has
peak probability at 1 and a relatively narrow width, due to the fact that it is
a tissue specific multiplicative modifying factor. To understand the roles of
these priors refer to (**SM3 Fig**).

Using this model, we split the data by Male/Female sex. Sex is known to be a
risk factor for death, with males dying at slightly higher rates for all ages of
interest in our dataset.

### 2021-06-30 text/survival.Rmd

We found that cancer types, as expected, have a range of prognoses relative to
the general population. In **Fig 5B-D** we show how effective age baseline, and
effective age rate (which modifies the baseline based on your actual age) as
well as the ktis which accelerates or decelerates further aging, varies
considerably from cancer to cancer. For example a 20 year old diagnosed with
THCA or PRAD would be threated similarly to a 40 year old based on teh Effective
Age parameter. It's noticeable that for every year beyond 20 in the PRAD
diagnosis we add only an extra ~ 0.2 years (Effective Age Rate). The ktis
parameter represents an acceleration (greater than 1) or deceleration (less) of
the further aging beyond diagnosis. For example for BRCA patients each year past
diagnosis adds about 1 year of risk, whereas for LGG patients it adds around 2,
and for SARC patients it adds only 0.25 or so.

>>>>

We found that cancer types, as expected, have a range of prognoses relative to
the general population. In **Fig 5B-D** we show how effective age baseline, and
effective age rate (which modifies the baseline based on your actual age) as
well as the $k_{tis}$ which accelerates or decelerates further aging, varies
considerably from cancer to cancer. For example a 20 year old diagnosed with
THCA or PRAD would be treated similarly to a 40 year old based on the Effective
Age parameter. It's noticeable that for every year beyond 20 in the PRAD
diagnosis we add only an extra ~ 0.2 years (Effective Age Rate). The $k_{tis}$
parameter represents an acceleration (greater than 1) or deceleration (less) of
the further aging beyond diagnosis. For example for BRCA patients each year past
diagnosis adds about 1 year of risk, whereas for LGG patients it adds around 2,
and for SARC patients it adds only 0.25 or so.
